---
title: "hw3"
author: 'IRENE MARTINEZ MORATA (UNI: im2557)'
date: "10/14/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(ggplot2)
library(p8105.datasets)
```

## Problem 1

There are 134 aisles in instacart


```{r load the data}
data("instacart")

instacart %>% 
  group_by(order_id) %>% 
  summarize(n = n())

instacart %>%
  group_by(product_id) %>%
  summarize(n = n())

instacart %>% 
  group_by(order_id) %>% 
  summarize( n_product = n_distinct(product_id))

instacart %>%
  group_by(user_id) %>%
  summarize(n = n())





```

How many aisles are there, and which aisles are the most items ordered from?

```{r problem 1 questions}

instacart %>%
  group_by(aisle) %>%
summarize(n = n()) %>%
arrange(desc(n))
  
  
instacart %>% 
  group_by(aisle_id, add_to_cart_order) %>% 
  count()
```
Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r }
plot1 <- instacart %>% 
  group_by(product_id, aisle) %>% 
  summarize(n = n()) %>%
  mutate(product_n = n >= 1000) %>% 
  filter(product_n == TRUE) %>%
  mutate(aisle = reorder(aisle, desc(n))) %>%
   ggplot(
       aes(x = aisle, y = n)) +
  geom_point() +
  labs(
    x = "aisle",
    y = "number of items"
  )  +
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
print(plot1)

```

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.


```{r }

instacart %>% 
  group_by(aisle, product_name) %>% 
  summarize(n = n()) %>% 
  filter(aisle == "baking ingredients" | aisle == "dog food care" | aisle == "package vegetables fruits") %>% 
  mutate(n_ranking = min_rank(desc(n)))
pivot_wider(
    names_from = aisle,
    values_from = product_name
  ) 
  
  mutate(n_ranking = min_rank(desc(n))) 
  
  select(aisle, product_name) %>% 
  pivot_wider(
    names_from = aisle,
    values_from = product_name
  )

```
